### 2019-04-30 report/implementation.tex
fiigures
Début de la vitesse de la lumière

>>>>

The result seems to be correct. We then try on larger images, the card images that we can find in the original paper presenting convolutional Wassertein \cite{Goe++}. Here is what we obtained:

\begin{figure}[h]
	\centering
	\begin{tikzpicture}
		\node[] at (0, 0) {\includegraphics[scale=0.3]{bar0.png}};
		\node[] at (2.8, 0) {\includegraphics[scale=0.3]{bar3.png}};
		\node[] at (5.6, 0) {\includegraphics[scale=0.3]{bar6.png}};
		\node[] at (8.4, 0) {\includegraphics[scale=0.3]{bar8.png}};
		\node[] at (11.2, 0) {\includegraphics[scale=0.3]{bar11.png}};
		\node[] at (14, 0) {\includegraphics[scale=0.3]{bar14.png}};
	\end{tikzpicture}
	\caption{Reproduction of \figurename~\ref{cardp}}
\end{figure}

\paragraph{SVM usage}
As proposed in the paper of Cuturi \cite{Cut}, we used the Wasserstein distance to compute the Gram matrix for 500 image and we used it as a kernel in a SVM. On a test set of 1500 images we obtained 92.4\% of good answers. Whereas with a kernel using Euclidean distance we obtain a good answer on only 88.4\%. To check these results you can run the python script \verb|learning.py| on the two kernels that are in the \verb|kernels| folder.

### 2019-04-30 report/overview.tex
Début de la vitesse de la lumière

\paragraph{Changement de plan de transport}
\paragraph{Change of transport plan}
The idea of Kantorovich is to change the function $T : \Omega \rightarrow
\Omega$ into a probability $P$ on the product space $\mathcal{P}(\Omega \times
\Omega)$. On the upper example, Monge's formulation forbids to separate the
masse of size 150 into two separate masses. With Kantorovich's formulation,
this operation is accessible: the mass of $150$ is splitted into a mass of
$90$ and a mass of $60$, both sent to different places. In the discrete
case, we can represent what happens with a matrix.

\begin{center}
	\begin{tikzpicture}
		\draw[step=1, thick] (0, 0) grid (3, 3);
		\node at (0.5, 0.5) {\Large $p_{{\color{red}3}{\color{blue}A}}$};
		\node at (1.5, 0.5) {\Large $p_{{\color{red}3}{\color{blue}B}}$};
		\node at (2.5, 0.5) {\Large $p_{{\color{red}3}{\color{blue}C}}$};
		\node at (0.5, 1.5) {\Large $p_{{\color{red}2}{\color{blue}A}}$};
		\node at (1.5, 1.5) {\Large $p_{{\color{red}2}{\color{blue}B}}$};
		\node at (2.5, 1.5) {\Large $p_{{\color{red}2}{\color{blue}C}}$};
		\node at (0.5, 2.5) {\Large $p_{{\color{red}1}{\color{blue}A}}$};
		\node at (1.5, 2.5) {\Large $p_{{\color{red}1}{\color{blue}B}}$};
		\node at (2.5, 2.5) {\Large $p_{{\color{red}1}{\color{blue}C}}$};

@@

		\node at (5.5, 4.5) {\large Interpolation linéaire};

@@

		\node at (5.5, 4.5) {\large Interpolation de Wasserstein};

@@

We recall that the computation of the barycenter $\mu$ of the $N$ distributions
$\nu_i$ associated to the weights $\lambda_i$ boils down to the following
minimization problem.
$$ \min_{\mu \in \mathcal{P}(\Omega)} \sum_{i=1}^N \lambda_i W_p^p(\mu, \nu_i) $$
Now observe what the barycenter looks like with $4$ distributions:
\begin{center}
	\includegraphics[scale=0.4]{wasserstein_bar.png}
\end{center}

@@

\paragraph{Sinkhorn}
We have at our disposition an interative method to find this matrix $P_\gamma$. We use the condition $P_\gamma \in U(a,b)$:
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrr}
\text{diag}(u) K \text{diag}(v) \mathbbm{1}_m & = & a \\
\text{diag}(v) K^\trans \text{diag}(u) \mathbbm{1}_n & = & b
\end{array} \right. $$
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrr}
\text{diag}(u) K v & = & a \\
\text{diag}(v) K^\trans u & = & b
\end{array} \right. $$
Then, we introduce $\odot$ the coefficient-wise product:
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrr}
u \odot K v & = & a \\
v \odot K^\trans u & = & b
\end{array} \right. $$
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrc}
u & = & a \, / \, K v \\
v & = & b \, / \, K^\trans u
\end{array} \right. $$
Sinkhorn's algorithm is thus the following: \\
\begin{algorithm}[H]
	\caption{Sinkhorn}
	\Repeat{convergence of $u, v$}{
		$u \gets a \, / \, K v$ \;
		$v \gets b \, / \, K^\trans u$ \;
	}
\end{algorithm}

>>>>

\paragraph{Change of transport plan}
The idea of Kantorovich is to change the function $T : \Omega \rightarrow
\Omega$ into a probability $P$ on the product space $\mathcal{P}(\Omega \times
\Omega)$. On the upper example, Monge's formulation forbids to separate the
masse of size 150 into two separate masses. With Kantorovich's formulation,
this operation is accessible: the mass of $150$ is splitted into a mass of
$90$ and a mass of $60$, both sent to different places. In the discrete
case, we can represent what happens with a matrix.

\begin{center}
	\begin{tikzpicture}[scale=1.15]
		\draw[step=1, thick] (0, 0) grid (3, 3);
		\node at (0.5, 0.5) {\Large $p_{{\color{red}3}{\color{blue}A}}$};
		\node at (1.5, 0.5) {\Large $p_{{\color{red}3}{\color{blue}B}}$};
		\node at (2.5, 0.5) {\Large $p_{{\color{red}3}{\color{blue}C}}$};
		\node at (0.5, 1.5) {\Large $p_{{\color{red}2}{\color{blue}A}}$};
		\node at (1.5, 1.5) {\Large $p_{{\color{red}2}{\color{blue}B}}$};
		\node at (2.5, 1.5) {\Large $p_{{\color{red}2}{\color{blue}C}}$};
		\node at (0.5, 2.5) {\Large $p_{{\color{red}1}{\color{blue}A}}$};
		\node at (1.5, 2.5) {\Large $p_{{\color{red}1}{\color{blue}B}}$};
		\node at (2.5, 2.5) {\Large $p_{{\color{red}1}{\color{blue}C}}$};

@@

		\node at (5.5, 4.5) {\large Linear Interpolation};

@@

		\node at (5.5, 4.5) {\large Wasserstein Interpolation};

@@

We recall that the computation of the barycenter $\mu$ of the $N$ distributions
$\nu_i$ associated to the weights $\lambda_i$ boils down to the following
minimization problem.
$$ \min_{\mu \in \mathcal{P}(\Omega)} \sum_{i=1}^N \lambda_i W_p^p(\mu, \nu_i) $$
Now you can observe what the barycenter looks like with $4$ distributions in the \figurename~\ref{shp1}.
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.4]{wasserstein_bar.png}
	\captionsetup{justification=centering}
	\caption{Each image contains several barycenters of the four shapes that are in the corners}
	\label{shp1}
\end{figure}

@@

\paragraph{Sinkhorn}
We have at our disposition an interative method to find this matrix $P_\gamma$. We use the condition $P_\gamma \in U(a,b)$:
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrr}
\text{diag}(u) K \text{diag}(v) \mathbbm{1}_m & = & a \\
\text{diag}(v) K^\trans \text{diag}(u) \mathbbm{1}_n & = & b
\end{array} \right. $$
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrr}
\text{diag}(u) K v & = & a \\
\text{diag}(v) K^\trans u & = & b
\end{array} \right. $$
Then, we introduce $\odot$ the coefficient-wise product:
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrr}
u \odot K v & = & a \\
v \odot K^\trans u & = & b
\end{array} \right. $$
$$ P_\gamma \in U(a, b) \Leftrightarrow \left\{ \begin{array}{lrc}
u & = & a \, / \, K v \\
v & = & b \, / \, K^\trans u
\end{array} \right. $$
Sinkhorn's algorithm is thus the following: \\
\begin{algorithm}
	\caption{Sinkhorn}
	\Repeat{convergence of $u, v$}{
		$u \gets a \, / \, K v$ \;
		$v \gets b \, / \, K^\trans u$ \;
	}
\end{algorithm}

### 2019-04-30 report/papers.tex
début implémentation
Début de la vitesse de la lumière

\sect{Sinkhorn Distances: Lightspeed Computation of Optimal Transport}

\cite{Cut}

@@

\paragraph{Applications}
This has many applications in geometric domains. Applications that are given are:
\begin{itemize}
	\item Shape interpolation
		\begin{center}
		\includegraphics[scale=0.22]{shape_int.png}
		\end{center}
	\item BRDF design
	\item Color histogram manipulation
	\item Skeleton layout
	\item Soft maps
\end{itemize}
Finally here are two images of barycenters they have computed. The first one is an image where we find barycenters of four images that are in the corners. The second is an interpolation between two cards. As we tried to reproduce it in the implementation part we put them here :

\begin{center}
	\includegraphics[scale=0.25]{bars_paper.png} \\ \vspace{3mm}
	\includegraphics[scale=0.4]{cards_paper.png}
\end{center}

>>>>

\sect{Sinkhorn Distances: Lightspeed Computation of Optimal Transport \cite{Cut}}

In this paper written in 2013, Cuturi proposed one of the most currently used algorithm to compute Wasserstein distances. He smoothed the classic optimal transport problem thanks to an entropy-regularization and it leads to a new distance which has a faster solver. He also showed that using regularization gives better results on MNIST classification problem than using the classic Wasserstein distance or obviously the Euclidean distance. This paper is part of the search for adequate and easily computable distances in machine learning. Wasserstein distance is better than Euclidean distance in many cases of machine learning but because the complexity to compute it is $\mathcal{O}(n^3 \log n)$ for $n$ the size of the data Wasserstein distance is rarely used. That's why techniques to compute the Wasserstein distance had to be develop.

\paragraph{Idea}
The main ideas of this paper are the use of an entropic-regularization to simplify computation when using Sinkhorn iterative algorithm and idea of computing the distances between one point and a family of other points in the same time, to use matrix-matrix multiplications instead of matrix-vector multiplications. Thus the algorithm can be implemented on GPU architectures.

\paragraph{Regularization}
We recall that the goal of the optimal transport problem in the Kantorovich formulation is to find a joint distribution $P \in \R^{d \times d}$ where the marginals equal the two distributions for which we want to compute the distance. That is to say $P \mathbbm{1}_n = a$ and $P^\trans \mathbbm{1}_n = b$ when we compute the distance between $a$ and $b$. We call $U(a, b)$ the set of all possible values for $P$. By denoting the entropy by $H$ we can show that:
$$ H(P) \leqslant H(a) + H(b) $$
And $H(P) = H(a) + H(b)$ when $P = a b^\trans$. We then restrict $P$ to the new set:
$$ U_\alpha(a, b) = \{ P \in U(a, b) \, | \, \text{KL}(P || a b^\trans) \leqslant \alpha \} = \{ P \in U(a, b) \, | \, H(P) \geqslant H(a) + H(b) - \alpha \} $$
Two reasons are given to this restriction. The first one is that it will be easier to compute the distance. The second is that without regularization $P$ has almost $2n-1$ non-zero coefficients. The transport is almost deterministic and it is not natural. That's why smoothing the transport plan with entropic regularization is presented as a good idea. In the paper, it is shown that this restriction induce a new distance:
$$ d_{M, \alpha} = \min_{P \in U_\alpha(a, b)} \langle P, M \rangle $$
Were $M$ is a distance matrix.

\paragraph{Computation}
By duality theory he obtained that to $\alpha$, corresponds a value $\lambda \geqslant 0$ such that the distance $d_{M, \alpha}$ is equal to the distance $d_M^\lambda$ defined by:
$$ d_M^\lambda(a, b) = \langle P^\lambda, M \rangle, \quad \text{where } P^\lambda = \argmin_{P \in U(a, b)} \langle P, M \rangle - \dfrac{1}{\lambda} H(P) $$
Then Cuturi show that this new distance can be computed with a much cheaper cost than the original Wassertein distance.

@@

\paragraph{Applications}
This has many applications in geometric domains. Applications that are given are:
\begin{itemize}
	\item Shape interpolation (\figurename~\ref{shape})
		\begin{figure}[h]
			\centering
			\includegraphics[scale=0.22]{shape_int.png}
			\caption{Application to shape interpolation}
			\label{shape}
		\end{figure}
	\item BRDF design
	\item Color histogram manipulation
	\item Skeleton layout
	\item Soft maps
\end{itemize}
Finally here are two images of barycenters they have computed. The first one is an image where we find barycenters of four images that are in the corners. The second is an interpolation between two cards. As we tried to reproduce it in the implementation part we put them here :

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.28]{bars_paper.png}
	\caption{Barycenters of four images}
	\label{barp}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[scale=0.45]{cards_paper.png}
	\caption{Interpolation of two images of cards for times $t = 0, 0.25, 0.5, 0.75, 1$.}
	\label{cardp}
\end{figure}
